---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidymapgrid

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tidymapgrid)](https://CRAN.R-project.org/package=tidymapgrid)
[![R-CMD-check](https://github.com/ian-flores/tidymapgrid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ian-flores/tidymapgrid/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/ian-flores/tidymapgrid/branch/main/graph/badge.svg)](https://codecov.io/gh/ian-flores/tidymapgrid?branch=main)
<!-- badges: end -->

tidymapgrid provides a simple interface for converting geographic polygons into grid representations (hexagonal or square). This is useful for creating tile grid maps where each geographic unit is represented by a uniformly-sized cell, making it easier to visualize data without the distortion caused by varying polygon sizes.

The package wraps the [geogrid](https://github.com/jbaileyh/geogrid) package and returns tidy sf objects that integrate seamlessly with the tidyverse and ggplot2.

## Installation

You can install the released version of tidymapgrid from [GitHub](https://github.com/ian-flores) with:

``` r
devtools::install_github('ian-flores/tidymapgrid')
```

## Example

This is a basic example which shows you how to obtain the grid for your map.

```{r library, warning=FALSE, message=FALSE}
library(sf)
library(tidymapgrid)
library(ggplot2)
```

```{r reg-ggplot, warning=FALSE, message=FALSE}
json_url <- 'https://raw.githubusercontent.com/ian-flores/Hurricane_Maria_Mortality_Analysis/master/analysis/data/pueblos.json'

pueblos <- st_read(json_url)

ggplot(pueblos) +
  geom_sf() +
  theme_void()
```

```{r get_grid, warning=FALSE, message=FALSE}
grid <- get_grid(pueblos, type = 'hexagonal', seed = 40)

ggplot(grid) +
  geom_sf() +
  theme_void()
```
