---
title: "Getting Started with tidymapgrid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with tidymapgrid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `tidymapgrid` package provides a simple interface for converting geographic
polygons into grid representations. This is particularly useful for creating
**tile grid maps** (also known as grid cartograms) where each geographic unit
is represented by a uniformly-sized cell.

### Why use grid maps?

Traditional geographic maps can be misleading because:

- **Large areas dominate visually** even when they're not important for the data
- **Small areas are hard to see** even when they contain important information
- **Irregular shapes** make it difficult to compare values across regions

Grid maps solve these problems by representing each geographic unit with an
equal-sized cell, making comparisons fair and easy.

## Installation

You can install tidymapgrid from GitHub:
```r
devtools::install_github("ian-flores/tidymapgrid")
```

## Basic Usage

```{r setup, message=FALSE, warning=FALSE}
library(tidymapgrid)
library(sf)
library(ggplot2)
```

### Loading geographic data

First, load your geographic data as an `sf` object. For this example, we'll
use a sample dataset of Puerto Rico municipalities that comes bundled with
the package:

```{r load-data, message=FALSE, warning=FALSE}
# Load sample data bundled with the package
pueblos <- readRDS(system.file("extdata", "pueblos.RDS", package = "tidymapgrid"))

# View original map
ggplot(pueblos) +
  geom_sf(fill = "steelblue", color = "white") +
  theme_void() +
  labs(title = "Puerto Rico Municipalities (Original)")
```

### Creating a hexagonal grid

The `get_grid()` function converts your polygons to a grid:

```{r hex-grid, message=FALSE, warning=FALSE}
# Create hexagonal grid
hex_grid <- get_grid(pueblos, type = "hexagonal", seed = 42)

ggplot(hex_grid) +
  geom_sf(fill = "steelblue", color = "white") +
  theme_void() +
  labs(title = "Puerto Rico Municipalities (Hexagonal Grid)")
```

### Creating a square grid

You can also create square grids:

```{r square-grid, message=FALSE, warning=FALSE}
# Create square grid
square_grid <- get_grid(pueblos, type = "square", seed = 42)

ggplot(square_grid) +
  geom_sf(fill = "coral", color = "white") +
  theme_void() +
  labs(title = "Puerto Rico Municipalities (Square Grid)")
```

## Working with Data

The grid retains all attributes from the original data, making it easy to
create thematic maps:
```{r thematic, message=FALSE, warning=FALSE}
# Add some example data
hex_grid$value <- runif(nrow(hex_grid), 0, 100)

ggplot(hex_grid) +
  geom_sf(aes(fill = value), color = "white") +
  scale_fill_viridis_c() +
  theme_void() +
  labs(
    title = "Thematic Grid Map",
    fill = "Value"
  )
```

## Reproducibility

The `seed` parameter ensures reproducibility. The same seed will always produce
the same grid layout:

```{r reproducibility}
grid1 <- get_grid(pueblos, type = "hexagonal", seed = 123)
grid2 <- get_grid(pueblos, type = "hexagonal", seed = 123)

# These are identical
identical(grid1, grid2)
```

Different seeds produce different arrangements, so you can experiment to find
the most pleasing layout for your data.

## Tips and Best Practices

1. **Experiment with seeds**: Try different seed values to find a layout that
   best preserves the geographic relationships in your data.

2. **Choose the right grid type**:
   - **Hexagonal** grids look more organic and are often preferred for
     statistical visualization
   - **Square** grids are simpler and may work better for certain applications

3. **Consider your data**: Grid maps work best when comparing values across
   regions. If the actual geography is important, a traditional map may be more
   appropriate.

## See Also

- The [geogrid](https://github.com/jbaileyh/geogrid) package, which provides
  the underlying algorithm
- [sf](https://r-spatial.github.io/sf/) for working with spatial data in R
- [ggplot2](https://ggplot2.tidyverse.org/) for visualization
